### 2013-08-26 text/tex/appendix.tex

\appendix

\section{More on Lensing Theory} \label{sec:more-theory}

First we define comoving distances, as follows.
\begin{equation}
\int \frac{dz}{\sqrt{\Omega_m(1+z)^3 + \Omega_\Lambda}}
\end{equation}
\begin{equation}
\begin{aligned}
&D_S    &                                &\int_0^{a_S} \\
\noalign{\medskip}
&D_{LS} & = \frac c{H_0} \ \times \quad  &\int_{z_L}^{z_S} \\
\noalign{\medskip}
&D_L    &                                &\int_0^{z_L}
\end{aligned}
\end{equation}

\begin{equation}
\begin{aligned}
A           &= cD_L \frac{D_{LS}}{D_S} \times t \\
\kappa(x,y) &= \frac{4\pi G}{c^2} \frac{D_L}{1+z} \frac{D_{LS}}{D_S}
               \times \Sigma(x,y)
\end{aligned}
\end{equation}

Let $x,y$ be physical distances on the lens plane and $D_L$ be
comoving distances.  Then
\begin{equation}
t_{\rm geom} = \frac{(1+z_L)^2}{2cD_L} \frac{D_S}{D_{LS}}
\left( (x-s_x)^2 + (y-s_y)^2 \right)
\end{equation}
\begin{equation}
\nabla^2 t_{\rm grav} = (1+z_L)\frac{8\pi G}{c^3} \, \Sigma(x,y)
\end{equation}

Can replace comoving distances $D$ with angular-diameter distances
$d$,
\begin{equation}
(x,y) = d_L (\theta_x,\theta_y)
\end{equation}
and so on.
\begin{equation}
\begin{aligned}
D_S &= (1+z_S) \, d_S \\
D_{LS} &= \frac{1+z_S}{1+z_L} \, d_{LS} \\
D_L &= (1+z_L) \, d_L
\end{aligned}
\end{equation}
Angular-diameter distances are always smaller than comoving distances.

Recover equations (2.1) to (2.6) from \cite{1986ApJ...310..568B}.