### 2013-11-25 text/tex/outlook.tex

\section{Extensions needed} \label{sec:todo}

As seen in \secref{tests.t1}, volunteers fail mostly in two situations:
when to identify an arclike structure while placing the points
and in 5 image configurations, that are quite symmetric, then to identify the correct ordering of the points.

The first can be improved by additional images from different filters (?)
Will be implemented with a next major \spl upgrade.

The second can be made less by better training (introduce the ruler), may be more filters help too?

We hope that those problems are also taken care of when staring collaborative modeling. (Already in progress)
Basics are available ('revise functionality') but better community tools would help (having overview, comparison between different models ect...)

Lenses with more than one maximum can be allowed for.  See Figure 5c
in \citep{2001ApJ...557..594R} and Figure 4b in
\cite{2003ApJ...590...39K}.

Supply several images from several bands.