### 2013-11-28 text/tex/stats.tex

\begin{table}\centering\begin{tabular}{llcc}\hline
 & & n & p \\
\hline
 N: & Total number of models & 119 & 1.00\\
\hline
 R1: & images approx. on right location & 110 & 0.92\\
 R2: & images with correct parity & 70 & 0.59\\
\hline
 E01: & inaccurate in arc & 21 & 0.18\\
 E02: & wrong parity in 3 lens conf. & 2 & 0.02\\
 E03: & identified 3 of 5 imgs. & 5 & 0.04\\
 E04: & modeled arc with single img. & 4 & 0.03\\
 E05: & $\pi$ rotated parity & 7 & 0.06\\
 E06: & $\pi$/2 rotated parity & 38 & 0.32\\
 E07: & missed faint img. & 1 & 0.01\\
 E08: & too many imgs in arc. & 5 & 0.04\\
 E09: & missed double img & 3 & 0.03\\
 E10: & too many imgs. & 1 & 0.01\\
\end{tabular}

\caption{1. col: id of model; 2.col: places of imgs idendited
  correctly; 3. col: more or less correct identification of extr
  points. 4.col: exact identification of extremal points. 5.col:
  type(s) of errors ocured....  1) inaccurate placement in an extended
  arc; 2) wrongly identified sad and min in 3 image configuration; 3)
  identified only 3 instead of 5 images; 4) tried to model an arc with
  a min instead of min-sad-min; 5) PI-err (rotation by 180 degrees; in
  5 image configuration, exchanged the ordering of the two saddle
  points); 6) PI/2-err (rotation by 90 degree; sad-min-sad-min-sad);
  7) missed faint image(s); 8) tried to model an arc with min-sad-min
  instead of only min; 9) did identify two close by images as one; 10)
  used 7 or more image to model a 5 image system. }

\label{tab:stats}
\end{table}