### 2013-12-12 text/tex/abstract.tex
more edits

\begin{abstract}
In Spacewarps **, volunteers are invited to search sky surveys for
lens candidates.  Here we report on a system that allows experienced
volunteers to model lens candidates.  A sample of 29 simulated lenses
were modeled, each multiple times.  The quality of the models was then
examined in different ways.  In particular: (i)~the image parities and
time orderings were identified correctly in most cases, but not
always; (ii)~the mean convergence (equivalent to the enclosed mass)
was quite well constrained in the image region; (iii)~the performance
of experienced volunteers was comparable to professionals.
\end{abstract}

\keywords{}