### 2014-07-27 text/ms.tex
Reordered authors to show properly Prasenjit's role, and to give the modellers full credit. Fixes #20. Also beefed up abstract a bit, addressing #19

% ====================================================================

\documentclass[usenatbib]{mn2e}

\usepackage{amsmath}
\usepackage{xspace}     % fixes spaces after custom commands
\usepackage{float}      % exact placement of figures (using \begin{figure}[H])
\usepackage[section,below]{placeins}   % gives a \FloatBarrier to force all floats to appear
\usepackage{subfigure}  % does what it says

\usepackage{graphicx}
% prefer pdf over png (better quality, lossless compression)
\DeclareGraphicsExtensions{%
  .pdf,.PDF,%
  .png,.PNG,%
  .jpg,.mps,.jpeg,.jbig2,.jb2,.JPG,.JPEG,.JBIG2,.JB2}

\newcommand{\apj}{ApJ}
\newcommand{\apjs}{ApJS}
\newcommand{\apjl}{ApJL}
\newcommand{\aap}{A{\&}A}
\newcommand{\aaps}{A{\&}AS}
\newcommand{\mnras}{MNRAS}
\newcommand{\aj}{AJ}
\newcommand{\araa}{ARAA}
\newcommand{\pasp}{PASP}
\newcommand{\pasj}{PASJ}
\newcommand{\nat}{Nature}
\newcommand{\nar}{New Astr Rev}

\def\half{{\textstyle\frac12}}
% makes it easier to have consistent writing, and easy change if needed
\newcommand{\spl}{SpaghettiLens\xspace}
\newcommand{\sw}{Spacewarps\xspace}

% shortcut for einstein radius (Text Greek Full Math)
\newcommand{\ERf}[1][]{$\Theta_\text{E#1}$\xspace} % full output

%\newcommand{\ER}{Einstein radius\xspace} % text
%\newcommand{\ERg}[1][]{$\Theta_{\text{E#1}}$\xspace} % er in textmode with greek
%\newcommand{\ERm}[1][]{\Theta_\text{E#1}} % mathmode with greek symbol

%shortcuts for kappa
%\newcommand{\kenc}[1][r]{$\kappa_\text{encl}(#1)$\xspace}
%\newcommand{\kap}[1][r]{$\kappa(#1)$\xspace}

% shorcuts for refs (use capital for beginning of sentence)
% first 3 are for the real layz people..
\newcommand{\fref}[1]{\ref{fig:#1}}
\newcommand{\sref}[1]{\ref{sec:#1}}
\newcommand{\tref}[1]{\ref{tab:#1}}
\newcommand{\figref}[1]{Figure~\ref{fig:#1}}
\newcommand{\secref}[1]{Section~\ref{sec:#1}}
\newcommand{\tabref}[1]{Table~\ref{tab:#1}}
\newcommand{\Figref}[1]{Figure~\ref{fig:#1}}
\newcommand{\Secref}[1]{Section~\ref{sec:#1}}
\newcommand{\Tabref}[1]{Table~\ref{tab:#1}}

% shortcut for ASW000xxxx
\newcommand{\asw}[1]{ASW000#1\xspace}
\newcommand{\model}[2][Model~]{#1#2\xspace}

%degrees
\newcommand{\dgr}{^{\circ}}

%more abbrevs
\newcommand{\tgeom}{t_{\rm geom}}
\newcommand{\tgrav}{t_{\rm grav}}
\newcommand{\subcirc}{{\lower.33ex\hbox{$\circ$}}}
\newcommand{\subbullet}{{\lower.33ex\hbox{$\bullet$}}}

% todo annotations on side in red. easy to search for at the end
% !!! REMOVE THIS LINES FOR FINAL COMPILE
%
\usepackage{setspace}    % remove this pgk for final compile!! 
\usepackage{color}    % well... color...

%\newcommand{\todo}[2][red]{%
%\textcolor{#1}{\textbullet}%
%\marginpar{\colorbox{#1}{\parbox{\marginparwidth}{%
%\setstretch{0.4}\sffamily\textcolor{black}{\scriptsize{#2}}}}}}

\newcommand{\needfig}[1][]{%
\colorbox{yellow}{\parbox{0.5\textwidth}{%
\vspace{0.5cm}\setstretch{0.5}\textcolor{black}{\scriptsize{fig #1}}\vspace{0.5cm}}}%
\marginpar{\colorbox{yellow}{\parbox{\marginparwidth}{%
\setstretch{0.4}\sffamily\textcolor{black}{\scriptsize{fig}}}}}}

\setlength{\marginparsep}{0mm}
\setlength{\marginparwidth}{2.2cm}

\newcommand{\needcite}[1][]{\todo[green]{cit #1}}
\newcommand{\needref}[1][]{\todo[cyan]{ref #1}}
\newcommand{\needchk}[1][]{\todo[red]{chk #1}}

\definecolor{purple}{RGB}{150,0,200}
\newcommand{\phil}[1]{\noindent\textcolor{purple}{\bf #1}}

% \\\ END OF REMOVAL SECION

\newlength{\myplotswidth}
\setlength{\myplotswidth}{0.4\linewidth}

\newcommand{\hr}{\vspace{5mm}\noindent\rule{0.8\textwidth}{0.4pt}\vspace{5mm}}

% ====================================================================

\begin{document}

\title{Gravitational Lens Modeling in a Citizen Science Context}

\author[Kueng et al]{Rafael Kueng,$^{1}$
Prasenjit Saha,$^{1}$
Elisabeth Baeten,$^{2}$
Claude Cornen,$^{2}$
\newauthor
Christine McMillan,$^{2}$
Julianne Wilcox,$^{2}$
Jonas Odermatt,$^{3}$
Jonathan Coles,$^{4}$
\newauthor
Aprajita Verma,$^{5}$
Anupreeta More,$^{6}$
Surhud More,$^{6}$
and Phil Marshall$^{7}$ \\
$^{1}$Physik-Institut, University of Zurich, 190 Winterthurerstrasse, 8057, Zurich, Switzerland\\
$^{2}$The Space Warps Collaboration\\
$^{3}$Kantonsschule Zug, L\"ussiweg 24, 6300 Zug, Switzerland\\
$^{4}$Exascale Computing Lab, Versailles Saint-Quentin-en-Yveline University, 55 Avenue des Etats Unis, 78000, Versailles, France\\
$^{5}$Oxford Astrophysics, Keble Road, Oxford OX1 3RH, UK\\
$^{6}$Kavli Institute for the Physics and Mathematics of the Universe, University of Tokyo, 5-1-5 Kashiwanoha, Kashiwa-shi 277-8583, Japan\\
$^{7}$Kavli Institute for Particle Astrophysics and Cosmology, Stanford University, 452 Lomita Mall, Stanford, CA 94035, USA\\
}

\maketitle

\begin{abstract}

We develop a method to enable collaborative modelling of gravitational
lenses, and lens candidates, by non-professional lens enthusiasts.  It
uses an existing free-form modelling program (GLASS), but
\phil{enables the input to this code to be provided in a novel way,}
via a user-provided diagram that is essentially a sketch of an
arrival-time surface.

An implementation \phil{of this method, \spl} is tested in a modelling
challenge using 29 simulated lenses \phil{drawn from the} set created
for \phil{the \sw citizen-science lens search}.  We find that
\phil{volunteers from this online community} asserted the image
parities and time ordering consistently in some lenses, but made
errors in other lenses depending on the image morphology. \phil{While}
errors in image parity and time ordering lead to  large errors in the
mass distribution, the enclosed mass \phil{was found to be more
robust: the model-derived Einstein radii found by the volunteers were
consistent with those produced by one of the professional team,
suggesting that given the appropriate tools, gravitational lens
modelling is a data analysis activity that can be crowd-sourced.}
Shortcomings \phil{of the implementation itself include} (a)~lens
profiles that tend to be systematically shallower than the lenses
\phil{(leading to systematic over-estimation of the Einstein radius 
by some $\sim25\%$}, and (b)~lack of detailed \phil{arc} modelling. 
Ideas for improvement are discussed.

\end{abstract}

\begin{keywords}
\end{keywords}

% ====================================================================

\section{Introduction}

The first two discoveries of gravitational lenses
\citep{1979Natur.279..381W,1980Natur.285..641W}, where a galaxy images
a background quasar into two (four), brought in their wake the first
work on lens modelling
\citep{1981ApJ...244..723Y,1981ApJ...244..736Y}.  For the second lens
(PG1115+080), mass models scored an early success with the prediction
that one of the lensed images seen would split further into two at
higher resolution.  That galaxies must sometimes cause multiple images
had long been expected \citep{1937ApJ....86..217Z}, and it had even
been argued that the phenomenon could help measure cosmological
parameters \citep{1964MNRAS.128..307R,1966MNRAS.132..101R}, but
apparently nobody was expecting that lenses would need detailed
modelling.  The first observations, however, immediately stimulated
models.  The reason for that lies in the image separation. Recall that
image separations are of order the angular Einstein radius
\begin{equation}
\sim \left(\frac{4GM}{c^2 D_L}\right)^{1/2}
\simeq 0.1'' \left(\frac{M}{M_\odot}\right)^{1/2}
             \left(\frac{D_L}{\rm pc}\right)^{-1/2}
\end{equation}
where $D_L$ is the distance to the lens, and $M$ its mass.  A lensing
galaxy with $M\sim10^{11}M_\odot$ at $10^9\rm\,pc$ would cause image
separations of $\sim1''$, which is comparable to the size of the
galaxy; typically the lensed images are seen through the galaxy halo.
Hence the lensed images depend on the detailed mass distribution of
the lensing galaxy.  Galaxy lenses demand modelling.

Since the early discoveries, more than 400 secure lenses are now known
(see Figure \ref{fig:masterlens}).  Modelling of the mass distribution
is part of any research using lenses, but so far no modelling study
has spanned all known lenses.  The largest single one
\citep{2009ApJ...703L..51K} models 58 lenses to infer the distribution
of dark matter around galaxies.  In other work,
\cite{2011ApJ...740...97L} combined lens models of 21 galaxies with
models of their stellar populations, to find the relation between
stars and dark matter, and \cite{2014MNRAS.437..600S} modelled 18
time-delay lenses together to infer cosmological parameters.

\begin{figure}
\centering
\includegraphics[width=\hsize]{fig/lenssky}
\caption{Sky distribution of 423 published lenses considered secure
  (from the Masterlens catalog at the University of Utah, maintained
  by Joel Brownstein and Leonidas Moustakas).  The map is in
  Hammer-Aitoff projection, with North up and $\rm RA=0$ in the
  middle. The empty swathes to the left and right of centre are the
  Milky Way. {\em Should we drop this figure?  It's not relevant to
    modelling, but as there seems to be no publicly available summary
    of all known lenses, it does help set context.}}
\label{fig:masterlens}
\end{figure}

Surveys now under way aim to increase the inventory of lenses another
ten or a hundred fold.  Among these is \sw (Marshall et al, in prep,
More et al, in prep).  \sw is a citizen science project\footnote{\tt
  http://www.spacewarps.org} in which volunteers are presented
sky-survey images and invited to identify lens candidates.  Simulated
lenses are mixed in with the data, both to help train volunteers on
what to look for, and to provide measures of the effectiveness of the
search.  The motivation for \sw is to enable volunteers, some of whom
had previously serendipitously identified lens candidates on earlier
citizen-science surveys, to make discoveries missed in automatic
searches by software robots.  Robots are very good with clean lensing
system in uncrowded fields with high signal-to-noise, but in general
test situations \citep{2009ApJ...694..924M} robots miss lenses (low
completeness) or contaminate the results with non-lenses (low purity).
\sw introduced its first tranche of survey data, in May 2013, from the
Canada-France-Hawaii Telescope Legacy Survey, or CFHTLS
\citep{Coupon2009}, covering $\simeq172$ square degrees (0.4\% of the
sky) was divided into tiles of $\simeq 80''\times80''$ ($440\times
440$ pixels).  Each such tile was shown to ten volunteers.  Despite
the rarity of detectable lenses ($<10^{-3}$ per tile), the \sw
collective soon found lens candidates, both rediscoveries already
known from robots \citep{Gavazzi2012, More2012ApJ} and possible new
lenses.

The encouraging early results prompted the question: could modelling
of the lenses also be done by volunteers?  If so, modelling could help
assess prioritise lens candidates at an early stage, which would be
very useful with new wide-field and sensitive surveys, which will
yield thousands of lens candidates. There are several software tools
for lensing modelling available, and work has been done on generic
interfaces \citep{2014A&C.....5...28L}.  Some early designs for \sw
included a prototype lens modelling tool \citep{2010AAS...21543527N}.
Moreover, some {\em Spacewarps\/} volunteers are quite experienced
from earlier projects, having spent a thousand hours or more with
data, and are very interested in more demanding projects.  The
interests of citizen-science communities are just beginning to be
studied \citep[e.g.,][]{2013AEdRv..12a0106J}, but it is clear that
some volunteers welcome open-ended challenges, and sometimes these
have led to new scientific results: one example is the discovery of an
exceptional extra-solar planet \citep{2013ApJ...768..127S}; another is
the development of new algorithms for protein folding
\citep{Khatib22112011}.  All these are grounds for optimism.  There
is, however, a basic difficulty in strong gravitational lensing.
Lensed images do not look much like their source, and still less do
they resemble the lensing-mass distribution.  To model a lens, one
needs either to do a lot of random guessing, or to have a good
intuition for what works.

In this paper, we propose a way around the difficulty, and report on a
modelling test on \sw using simulated lenses.  The three following
sections are devoted to the concept, the implementation, and tests
respectively.

In \secref{Fermat} we introduce a markup system for lensed images,
which we call a ``spaghetti diagram''.  A spaghetti diagram resembles
the visible image system, in a cartoon-like way, and at the same time
it encodes the basis of a mass model.  This supplies an intuitive link
between the image system and the mass distribution, which look
frustratingly different from each other.  Spaghetti diagrams are
essentially the saddle-point contours originally introduced to
gravitational lensing by \cite{1986ApJ...310..568B} as a way of
classifying lensed images.  They are sometimes shown as part of the
output of lens models \citep[for
  example][]{2001ApJ...557..594R,2003ApJ...590...39K,Lubini2012}.  In
the present work, however, spaghetti diagrams are the {\em input\/}
through which the modeller tells the program what to do.

In \secref{SpaghettiLens} we describe the \spl program, which
implements everything.  \spl is an extension of the GLASS framework
for modelling lenses \citep{2014arXiv1401.7990C}.  We will not go into
software details in this paper, concentrating instead on lens
modelling per se, but we remark that \spl is designed to be friendly
to the forum style of citizen-science projects, and enable incremental
model refinement by different people, without sacrificing any features
of GLASS.

In \secref{mod_challenge} we describe a modelling challenge where a
diverse sample of 29 simulated lenses was modelled multiple times
using \spl.  The models are then examined in two ways.  One was
whether the spaghetti diagram was correct.  The other was the recovery
of the Einstein radius.  In addition, we show some visual comparisons
of actual and recovered lens shape and radial profile, and identify
some areas to improve, but do not compute statistics on these aspects.
Profile and shape recovery with GLASS has been studied in more detail
in \citep{2014arXiv1401.7990C}.

\secref{todo} gives the general outlook and next steps.

% ====================================================================

\section{Fermat's principle and spaghetti diagrams} \label{sec:Fermat}

We now explain the lensing theory relevant to \spl, following the
formulation of gravitational lensing in terms of Fermat's principle by
\cite{1986ApJ...310..568B}.

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\subsection{Geometrical and gravitational time delays}

Consider a lens at some redshift $z_L$ and let $(x,y)$ be planar
coordinates at the lens, transverse to the line of sight.  Let
$\Sigma(x,y)$ be the mass per unit area, i.e., density projected along
the line of sight.  This surface density is often given in a
dimensionless form
\begin{equation} \label{eq:kappa}
\kappa(x,y) \propto \Sigma(x,y)
\end{equation}
called the convergence.  And let there be light, in the form of a more
distant source, at redshift $z_S$, behind point $(x_s,y_s)$ on the
lens.

We now imagine a virtual photon flying from the source to some $(x,y)$
on the lens, then changing direction and coming to the observer.  Such
a direction change would increase the light travel time compared to
coming through $(x_s,y_s)$.  The increased light travel time from the
geometry of deflection would be
\begin{equation} \label{eq:tgeom}
\tgeom(x,y) \propto (x-x_s)^2 + (y-s_y)^2 \,.
\end{equation}
assuming the delay is small compared to the total light travel time.

An additional delay of the photon comes from travelling through the
curved spacetime at the lens.  This gravitational time delay $\tgrav$
is related to the mass distribution of the lens.  The relation is
generally written as a two-dimensional Poisson equation, but an
alternative expression, avoiding calculus, is as follows.  The value
of $\tgrav$ through $(x,y)$ equals its average value on the
circumference of a small circle centred at $(x,y)$, plus a constant
times the mass within that circle.  The constant is $2G/c^3$ times the
cosmological expansion factor $(1+z_L)$. Thus
\begin{equation} \label{eq:tgrav}
\tgrav(x,y) = \left\langle \tgrav(x_\subcirc,y_\subcirc) \right\rangle
              + (1+z_L) \frac{2G}{c^3} M(x_\subbullet,y_\subbullet) \,.
\end{equation}
We have used $(x_\subcirc,y_\subcirc)$ to denote the circumference of
a circle, and $(x_\subbullet,y_\subbullet)$ to indicate the integrated
mass within the circle.

The light travel time of a virtual photon is therefore longer by
\begin{equation}  \label{eq:tarriv}
t(x,y) = t_{\rm geom} + t_{\rm grav}
\end{equation}
than it would have been with no lens present.  Real photons take paths
that make $t(x,y)$ extremal, that is, having a minimum, maximum or
saddle point (Fermat's principle).

The proportionality factors in \eqref{eq:kappa} and \eqref{eq:tgeom}
depend on the redshifts and cosmological parameters, and are given in
Appendix~\ref{more-theory}.

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{fig/arriv_0}
\includegraphics[width=\columnwidth]{fig/arriv_1}
\includegraphics[width=\columnwidth]{fig/arriv_2}
\caption{Perspective views and contour maps of example arrival time
  surfaces.  Contours are coloured in rainbow order (red: least delay,
  violet: highest delay).  The special contours that self-cross at
  saddle points are the basis of spaphetti diagrams.  {\em Upper
    panel:\/} No lens, hence showing the parabolic shape of the
  geometrical time delay.  The image would be at the bottom,
  coinciding with the source. {\em Middle panel:\/} A circular lensing
  mass (offset from the source) has been added, which has pushed the
  minimum to one side and introduced a maximum and saddle point, each
  corresponding to an image.  The saddle-point is characterized by a
  self-crossing ``spaghetti'' contour. {\em Lower panel:\/} An
  elongated lensing mass has been added.  There are now two minima,
  two saddle points, and a maximum, each corresponding to an image.}
\label{fig:arriv}
\end{figure}

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\subsection{Arrival-time contours} \label{sec:arriv}

The full function $t(x,y)$, also known as the arrival-time surface,
applies to virtual photons.  In other words, it is an abstract
construct and not itself observable.  But observable image positions
can be derived from the arrival-time surface, so visualising the
surface is useful.  \figref{arriv} does so.  In this figure, a
maximum, if present, is easy to see.  To locate mimima and
saddle-points, however, one needs to examine the contours of equal
arrival time.  A saddle point is characterised by a contour crossing
itself, forming an~X.  Mimima, on the other hand, have contours
looping around them, as do maxima.

The saddle-point contours which form an X are especially interesting,
because they set the overall topography of the arrival-time surface.
The obviously give the locations of the saddle points, and roughly
localise the minima and maxima as well.  If more precise locations of
the minima and maxima are added, the whole arrival-time surface is
already approximately known.  Since the arrival-time surface has an
exact relation to the lens-mass distribution and the source position,
in effect the mass distribution is also automatically approximately
specified.  In other words, a simple sketch of saddle-point contours
along with locations of minima and maxima ---which we call a
``spaghetti diagram''--- is implicitly already an approximation to a
lens-mass distribution.

The preceding assumes a point source.  To get an idea of what an
extended source would do, let us imagine moving the original source
slightly.  The contours of constant arrival time will naturally move
slightly, and so will the images.  The movement of the contours will
be most noticeable where the contours are far apart, that is where the
arrival-time surface is nearly flat.  As is evident from
\figref{arriv}, this is the region where the minimum and saddle points
lie, or near the images.  In this region, points on the source that
are close together produce images that are comparatively far apart.
In other words, the image is highly magnified.  In summary, lower
curvature in the arrival-time surface for a point source implies
larger magnification of an extended source.  Conversely, where the
arrival-time surface is strongly curved, the image will be
demagnified.  We see from \figref{arriv} that the arrival-time surface
tends to be highly curved near the maximum.  Hence maximum tend to be
demagnified.  In practice, maxima of the arrival time are nearly
always too faint to see. The minima and saddle points dominate.

% ====================================================================

\section{A lens-modelling program} \label{sec:SpaghettiLens}

\spl is not part of \sw, but it makes use of the \sw infrastructure,
in particular, the image database and the forum.  The forum is
essential essential for initial contact with interested members of the
\sw community.  Lens modelling from \sw works, in practice, in the
usual style of medium-sized astronomical collaborations, with
video-conferencing and in-person meetings where possible. But
preliminary results are immediately summarized on modelling threads on
the forum, and anyone interested is welcome to join at any time.
Another aspect of \sw (and of the Zooniverse family of projects in
general) that \spl adopts, is to keep all user interaction in a web
browser, with intensive computing done on a server.  Several
interesting design and implementation issues arise from these
considerations, but we will not go into them in this paper,
concentrating instead on aspects specific to lens modelling.

Modelling with \spl involves three stages, markup of the image,
followed by intensive numerics in the background, followed by review
of diagnostics and possible discussion.  We now describe each of
these.

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\subsection{Image markup}

One begins by going to the \spl page and entering the number of a \sw
image tile.  \spl presents the image, along with zoom and pan options
and a markup tool to construct a spaghetti diagram.  The human
modeller now has to make an educated guess for the topography of the
arrival-time surface, and input the locations time-ordering of the
maxima, minima, and saddle-points.  The markup tool \citep[which is
  inspired by Figure~6 of][and is like that figure made interactive
  and overlaid on data]{1986ApJ...310..568B} lets the modeller enter
the information by sketching saddle-point contours.  Examples can be
seen in the upper-left panels of Figures \ref{fig:6941} to
\ref{fig:7022}.  The loops in the markup tool were the origin of the
``spaghetti'' metaphor.

The markup tool allows only valid lensing configurations to be
entered.  The user does not need to think explicitly the image
parities (though the markup tool provides this information using
colour codes) or about time-ordering, or worry about the odd-image
theorem.  The exact placement of the loops in a spaghetti diagram has
no significance.  Only the heirarchy of which loop is inside which is
relevant.  The loops are there simply to help modeller's intuition.

As implemented so far, \spl assumes that the lens is dominated by a
single galaxy.  Only one maximum is permitted, and it is taken as the
centre of the main lensing galaxy.  The user can, however, mark
additional minor galaxies.  The latter are modelled as point
masses.

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\subsection{Numerics}

Having sketched a spaghetti diagram, the user presses a button to
initiate the next stage.  \spl then translates the spaghetti diagram
into input for GLASS, and forwards this input.  The task of GLASS,
which runs server-side as it is compute-intensive, is to find a mass
distribution $\kappa(x,y)$ that exactly reproduces the given locations
of the maximum, minima and saddle points. This criterion by itself is
extremely under-determined --- there are infinitely many mass
distributions that will reproduce a given set of maxima, minima and
saddle points, but typically they (a)~produce lots of extra images,
and (b)~look very unlike galaxies.  Additional assumptions (a prior)
are necessary.  GLASS uses the following prior
\citep[cf.][]{1997MNRAS.292..148S,2008ApJ...679...17C}.
\begin{enumerate}
\item The mass distribution is built out of non-negative tiles of
  mass.  (Sometimes these tiles are called mass pixels, but we should
  emphasize that they are unrelated to image pixels, and are much
  larger.)
\item There is a notional lens center, say $(x_0,y_0)$ which is
  identified with the maximum of the arrival time.  The source can
  have an arbitrary offset with respect to the lens center.
\item The mass distribution must be centrally concentrated, in two
  respects.  First, the circularly averaged density must fall away
  like $$ \left[(x-x_0)^2+(y-y_0)^2\right]^{-1/2}$$ or more steeply.
  Second, the direction of increasing density at any $(x,y)$ can point
  at most $45^\circ$ away from $(x_0,y_0)$.
\item The lens must be symmetrical with respect to $180^\circ$ rotations
  about $(x_0,y_0)$.  This symmetry assumption can be relaxed if the
  user wishes.
\end{enumerate}
There are still infinitely many models that satisfy both data and
prior constraints, but now they are more credible as galaxy lenses.
It is then possible to generate an ensemble of models.  The sampling
technique used by GLASS is described in \citep{Lubini2012}.
Typically, ensembles of 200 models are used.  That is to say, what we
call a \spl model is really the mean of an ensemble of 200 models, and
its estimated uncertainty is the range covered by the whole ensemble.

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\subsection{Diagnostics} \label{sec:diag}

After the model ensemble has been generated, \spl post-processes it to
present results and diagnostics to the user for inspection. This takes
the form of three figures.
\begin{enumerate}
\item A synthetic image of the lensed features.
\item A contour map of the arrival-time surface $t(x,y)$.
\item A gray scale plus contour map of the mass distribution.
\end{enumerate}
The synthetic image generated by \spl assumes a simple conical source
profile.  The user can change the contrast level on the image, which
(though it is not saved) amounts to adjusting the width of the cone.
These synthetic images are still very crude and not very useful for
assessing models.  The best indicator, in practice, of whether the
modelling was successful is contour map of $t(x,y)$, with saddle-point
contours highlighted.  It is, in effect, the computer's refinement of
the spaghetti diagram input by the user.  If the arrival-time surface
looks qualitatively similar to the spaghetti diagram, that generally
indicates a successful model.  The mass distribution also provides
indications; successful models generally lead to smooth-looking mass
distribution, whereas an irregular or checkboard pattern in the mass
map signals a bad model.

After examining this feedback, the user can choose to post the model
on the \spl archive.  They can also modify their input and try again,
or discard the attempt altogether.  After archiving, there can be
discussion among modellers, through the \sw forum or by any other
means, and revision of the model.  Any archived model can be revised
by any user: they can modify the spaghetti configuration slightly or
drastically, or change options like the size of the mass tiles.
Particularly interesting lens candidates lead to trees of models in
this way.  Discussion among modellers tends to prune a model tree,
focusing attention on the most interesting models.\footnote{See
  ``Collaborative gravitational lens modelling\dots'' in {\tt
    http://letters.zooniverse.org} for an example.}

% ====================================================================

\section{A lens modelling challenge} \label{sec:mod_challenge}

We now describe a test of the lens-modelling system, under conditions
that mimic as closely as possible the modelling of real lens
discoveries.  The lenses to be modelled were the simulated lenses
(known as ``sims'') already sprinkled onto the \sw field.  Once a
small user base had grown around \spl, a modelling challenge was
announced through the \sw forum.  The challenge set consisted of 29
sims, chosen to represent the different visual morphologies of \sw
sims. Modellers then contributed 119 models for these sims (at least
two for each sim).  Models were reported on the same forum used to
model real candidate lenses.  Modellers were free to consult and
refine each other's models, but had no information on how the sims
were generated.

Once the modelling was complete, the models were compared with the
originals.  There were two main tests: a check of whether the
spaghetti diagrams were correct for the lens in question, and a
comparison of the effective Einstein radii of the sims and the models.

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\subsection{The simulated lenses} \label{sec:sims}

The \sw sims are described in detail in More et al (in prep), but
relevant here is that the sims were of three kinds, as follows.

\begin{enumerate}
  \item Imitating lensed quasars: having a singular elliptical
    isothermal lens (SIE) plus constant external shear, and a circular
    Gaussian source.
  \item Emulating lensed galaxies: similar to the above, but with an
    elliptical de Vaucouleurs source.
  \item Resembling cluster lenses: having a source as above, but a
    more complicated lens, with one dominant elliptical SIE and
    one or more perturbing elliptical SIEs, plus a circular NFW
    \citep{1996ApJ...462..563N,1997ApJ...490..493N} to represent
    the underlying dark matter distribution.
\end{enumerate}
The {\tt gravlens} program \citep{2001astro.ph..2340K} was used.
Formulas for the lenses appear in \cite{2001astro.ph..2341K}. The SIE
lenses follow equations (33--35) of that work, with core radius set to
zero.  The NFW lens is in equations (48) and (50), while shear is the
$\gamma$ term in equation (76).

The information in this section was not revealed to the main developer
of \spl (RK, who also chose the challenge set) or to the modellers
(EB, CC, CM, JO, PS and JW) while modelling was in progress.  That is,
the modellers had no advance knowledge of what kind of
parameterisation had been used to make the sims.  After the modelling
stage, AM released the details of the sims for post-modelling
analysis.  Results from the latter now follow.

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\subsection{Some example models} \label{sec:example_models}

Of the 119 models proposed, we now discuss eight examples in some
detail.  Results from these are shown in Figures
\ref{fig:6941}--\ref{fig:7022}.  The first four of these show the most
common image morphologies, the other four explain some problem cases.

Each of Figures \ref{fig:6941}--\ref{fig:7022} figures has the
following layout.
$$ \begin{matrix}
\hbox{marked-up image} \qquad &\hbox{model synthetic image} \\
t(x,y)                        &\hbox{model } t(x,y) \\
\kappa(x,y)                   &\hbox{model } \kappa(x,y)
\end{matrix} $$
The marked-up image is a zoom of the lensed image on Spacewarps,
marked up with one or more spaghetti contours; this is the modeller's
input to \spl.  The three panels on the right show the graphical
output returned to the modeller by \spl, listed as (i), (ii), (iii) in
\secref{SpaghettiLens}.  These derive from the mean of an ensemble of
200 models generated by \spl. The model synthetic image is too crude
to be useful, because there is no fitting for the source light
distribution, as already noted in \secref{diag}.  The quality of the
synthetic images can be improved by interpolation, but as this feature
was not available during the modelling challenge, we do not show the
improved versions here.

\FloatBarrier

\begin{figure}
  \centering
  \includegraphics[width=\myplotswidth]{fig/006941_input}
  \includegraphics[width=\myplotswidth]{fig/006941_arr_time_ipol} \\
  \includegraphics[width=\myplotswidth]{fig/ASW000102p_006941_arriv} 
  \includegraphics[width=\myplotswidth]{fig/006941_spaghetti} \\
  \includegraphics[width=\myplotswidth]{fig/ASW000102p_006941_kappa}
  \includegraphics[width=\myplotswidth]{fig/006941_mass}
  \caption[result 6941 (ASW000102p)]{A simulated lens that mimics a
    lensed quasar, and model results.  The left panels derive from the
    simulation, and the right panels are \spl output.  Details of
    individual panels are in \secref{example_models}.}

  \label{fig:6941}
\end{figure}

\begin{figure}
  \centering

  \includegraphics[width=\myplotswidth]{fig/006990_input}
  \includegraphics[width=\myplotswidth]{fig/006990_arr_time_ipol} \\
  \includegraphics[width=\myplotswidth]{fig/ASW0004oux_006990_arriv}
  \includegraphics[width=\myplotswidth]{fig/006990_spaghetti} \\
  \includegraphics[width=\myplotswidth]{fig/ASW0004oux_006990_kappa}
  \includegraphics[width=\myplotswidth]{fig/006990_mass}

  \caption[result 6990 (ASW0004oux)]{Results from a system with an arc
    plus a counter-image, typical of lensed galaxies. (See Section
    \ref{sec:example_models} for details.)}
  \label{fig:6990}
\end{figure}

\begin{figure}
  \centering

  \includegraphics[width=\myplotswidth]{fig/006919_input}
  \includegraphics[width=\myplotswidth]{fig/006919_arr_time_ipol} \\
  \includegraphics[width=\myplotswidth]{fig/ASW0002z6f_006919_arriv}
  \includegraphics[width=\myplotswidth]{fig/006919_spaghetti} \\
  \includegraphics[width=\myplotswidth]{fig/ASW0002z6f_006919_kappa}
  \includegraphics[width=\myplotswidth]{fig/006919_mass}

  \caption[result 6919 (ASW0002z6f)]{Another configuration of arc plus
    counter-image; here the arc is closer to the lensing galaxy than
    the counter-image. (See \secref{example_models} for details.)}
  \label{fig:6919}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\myplotswidth]{fig/006915_input}
  \includegraphics[width=\myplotswidth]{fig/006915_arr_time_ipol} \\
  \includegraphics[width=\myplotswidth]{fig/ASW0001hpf_006915_arriv}
  \includegraphics[width=\myplotswidth]{fig/006915_spaghetti} \\
  \includegraphics[width=\myplotswidth]{fig/ASW0001hpf_006915_kappa}
  \includegraphics[width=\myplotswidth]{fig/006915_mass}

  \caption[result 6915 (ASW0001hpf)]{A four-image configuration
    typical of lensed quasars. (See \secref{example_models} for
    details.)}
  \label{fig:6915}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\myplotswidth]{fig/006975_input}
  \includegraphics[width=\myplotswidth]{fig/006975_arr_time_ipol} \\
  \includegraphics[width=\myplotswidth]{fig/ASW000195x_006975_arriv}
  \includegraphics[width=\myplotswidth]{fig/006975_spaghetti} \\
  \includegraphics[width=\myplotswidth]{fig/ASW000195x_006975_kappa}
  \includegraphics[width=\myplotswidth]{fig/006975_mass}
  \caption[result 6975 (ASW000195x)]{A lens with unrecovered mass
    substructure. (See \secref{example_models} for details.)}
  \label{fig:6975}
\end{figure}

\begin{figure}
  \centering

  \includegraphics[width=\myplotswidth]{fig/006937_input}
  \includegraphics[width=\myplotswidth]{fig/006937_arr_time_ipol} \\
  \includegraphics[width=\myplotswidth]{fig/ASW0000vqg_006937_arriv}
  \includegraphics[width=\myplotswidth]{fig/006937_spaghetti} \\
  \includegraphics[width=\myplotswidth]{fig/ASW0000vqg_006937_kappa}
  \includegraphics[width=\myplotswidth]{fig/006937_mass}

  \caption[result 6937 (ASW0000vqg)]{A sim with unrecovered
    substructure, resulting in a poor mass model. (See Section
    \ref{sec:example_models} for details.)}
  \label{fig:6937}
\end{figure}

\begin{figure}
  \centering

  \includegraphics[width=\myplotswidth]{fig/007025_input}
  \includegraphics[width=\myplotswidth]{fig/007025_arr_time_ipol} \\
  \includegraphics[width=\myplotswidth]{fig/ASW0000h2m_007025_arriv}
  \includegraphics[width=\myplotswidth]{fig/007025_spaghetti} \\
  \includegraphics[width=\myplotswidth]{fig/ASW0000h2m_007025_kappa}
  \includegraphics[width=\myplotswidth]{fig/007025_mass}
  \caption[result 7025 (ASW0000h2m)]{A four-image system with image
    parities incorrectly identified.  The model is poor, but the
    estimated Einstein radius is not bad. (See Section
    \ref{sec:example_models} for details.)}
  \label{fig:7025}
\end{figure}

\begin{figure}
  \centering

  \includegraphics[width=\myplotswidth]{fig/007022_input}
  \includegraphics[width=\myplotswidth]{fig/007022_arr_time_ipol} \\
  \includegraphics[width=\myplotswidth]{fig/ASW0000h2m_007022_arriv}
  \includegraphics[width=\myplotswidth]{fig/007022_spaghetti} \\
  \includegraphics[width=\myplotswidth]{fig/ASW0000h2m_007022_kappa}
  \includegraphics[width=\myplotswidth]{fig/007022_mass}

  \caption[result 7022 (ASW0000h2m)]{The same system as in Figure
    \ref{fig:7025}, this time with image parities correctly
    identified. (See \secref{example_models} for details.)}
  \label{fig:7022}
\end{figure}

\FloatBarrier

In these figures, the scales and alignments in the six panels are not
completely consistent: the three panels on the right have the same
coordinates; the middle-left and lower-left panels are consistent with
each other, but only approximately aligned with the right column,
because the latter are affected by errors in mouse placements; the
upper-left panel is zoomed out slightly, to accommodate the full
markup.  For qualitative features, which is all that
Figures~\ref{fig:6941}--\ref{fig:7022} try to convey, the slightly
varying scales and alignments are harmless.

Let us now consider these cases in turn.

\begin{itemize}

\item Figure \ref{fig:6941} shows the simplest case, with two clear
  images produced by a nearly-circular lens.  The centre of the
  lensing galaxy is a maximum, the image nearer to the galaxy is a
  saddle point, and the image further away is a minimum.  All these
  were correctly identified.  As note above, in
  \secref{SpaghettiLens}, the precise shape of the loops in the
  spaghetti diagrams is unimportant, only the implied image locations,
  parities and time-ordering matters.

\item Figure \ref{fig:6990} shows an example of an arc that has split
  into three images.  This kind of configuration, with a counter-image
  close to the lensing galaxy and a more distant arc/triplet on the
  other side, generically arises from an elongated mass distribution
  when the source is displaced along the elongated direction.  The
  spaghetti diagram in this case has another markup element, a grey
  point and circle overlaid on a probable secondary lensing galaxy.
  This is an instruction to \spl to allow a point mass at that
  location, distinct from the main mass map.

\item Figure \ref{fig:6919} shows another example of an arc plus
  counter-image, but (in contrast to \figref{6990}) the arc is closer
  to the lens than the counter-image. This configuration arises if the
  source displacement is perpendicular to the long axis of the lensing
  mass.  Comparing the two panels in the middle row, we see that the
  modeller interpreted arc as consisting of three images, whereas the
  sim shows a single saddle point associated with the arc.  But the
  identification is not really erroneous --- we just need to take into
  account that the source is extended.  In fact, in the sim the
  brightest part of the source is only doubly imaged, but the source
  extends into a region that produces four images.  In the $t(x,y)$ of
  the sim, the hairpin-bending contours are typical of double on the
  verge of splitting into a quad.

\item Figure \ref{fig:6915} shows another quad.  This kind of
  configuration arises when the mass is elongated and the source is
  displaced at an angle to the elongation.  The minima and saddle
  points are correctly identified, and the orientation of the
  ellipticity of the mass distribution is correctly reproduced.

\item Figure \ref{fig:6975} shows a lens with substructure in the form
  of a smaller secondary galaxy.  The galaxies in such group or
  cluster sims were based on galaxies visible in the images, but the
  modelers were not told in advance whether this was the case.  The
  minimum and saddle point are correctly identified.  The mass
  distribution misses the substructure, but overall appears
  reasonable.

\item Figure \ref{fig:7025} shows a quad.  In this one, the
  identification of the minima and saddle points was incorrect, and
  mass distribution comes out elongated East-West instead of
  North-South.  The mass distribution also appears somewhat jagged and
  the saddle-point contours are not as clean as in the previous
  examples; these are often indicators of a problem with the model.
  The enclosed mass is, however, none the worse --- the reason is
  probably that in a relatively symmetrical image configuration, the
  Einstein radius is quite well constrained by the images in a fairly
  model-independent way.

\item Figure \ref{fig:7022} shows another model of the same system.

\end{itemize}

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\subsection{Test of image identification} \label{sec:tests.t1}

The first post-modelling test was a qualitative comparison of the
original arrival-time surfaces and the input spaghetti diagrams given
by the modeller.  This tested first, for correct identification and
location of the lensed images, and second, the correct parities and
ordering for the lensed images in respect of the arrival time.

While we expected the identification of lensed images to be trivial,
given the generally clean appearance of the sims in the test, we
expected the parities and time-ordering to be more difficult.  While
the \spl tutorials had provided some general rules of thumb, to be
consistently correct with the time-ordering, a modeller needs to
develop some intuition for arrival time surfaces.  This is an area
where experience and tutorials training could improve results at a
later stage, and correspondingly, feedback on the difficulties
modellers encounter can help improve the tutorial materials.

\tabref{stats} presents a summary of the test.  The evaluation was
done manually, comparing the input to \spl with the actual
arrival-time surface of the sim.  This amounts to comparing the
upper-left and middle-left panels in each of
Figures~\ref{fig:6941}--\ref{fig:7022}, and similarly the other 111
models.

\begin{table}\centering\begin{tabular}{lr}\hline
    Total & 119 \\
\hline
    errors in image locations  & 9 \\
    errors in image parities or time ordering & 49 \\
\hline
    inaccurate image placement over an arc & 21 \\
    identified two images of four & 5 \\
    identified two nearby images as one & 3 \\
    missed faint images & 1 \\
    proposed too many images & 1 \\
    modelled a three-image arc as one image & 4 \\
    modelled one image as a three-image arc & 5 \\
    swapped minimum and saddle in double  & 2 \\
    swapped minima and saddles in quad & 38 \\
    swapped early and late saddles in a quad & 7 \\
\hline
\end{tabular}
\caption{Table of image-identification errors and the number of models
  containing each.  A model can contain more than one type of error.}
\label{tab:stats}
\end{table}

The images of the system are considered to have been located
correctly, if all the images have been identified and are approximated
within about 5\% of \spl frame used to draw the spaghetti diagram.
That frame size is adjustable by the user, but in practice it is
somewhat larger than the spaghetti diagram.  Such image-placement
errors were found in only 9 models.  That does not include inaccurate
image placement over an arc, which was considered a separate category
of error.

In addition to simple image-placement errors, ten types of errors were
recognised and are listed in \tabref{stats}.  Most of the problems
where due to unclear arc-like structures.  Critical errors like the
failure to identify all five images in a five images system, or to
include too many images, were rare.

The assignment of the parity of the images was a more difficult task,
and was successful in only 60\% of the cases.  The most common error
was swapping of minima and saddle-points in a quad; \figref{7025}
shows an example.  Another, less common, error was flipping the
spaghetti diagram, this swapping the time-ordering of the two saddle
points.

Incorrect image parities and time orderings tended to produce
poorer-looking models, such as the checkboard patterns in the mass map
in \figref{7025}.  Interestingly, however, the enclosed-mass profiles
were quite robust.  We will consider this aspect in the next section.

\FloatBarrier

\begin{figure}
  \centering
  \subfigure{\includegraphics[width=0.8\linewidth]
             {fig/007025_kappa_encl}}
  \subfigure{\includegraphics[width=0.8\linewidth]
             {fig/007022_kappa_encl}}
  \subfigure{\includegraphics[width=0.8\linewidth]
             {fig/007021_kappa_encl}} \\
  \caption{Mean $\kappa$ inside a circle around the lens centre, as a
    function of the radius of the circle.  (See \secref{tests.t2} for
    details.) The upper panel corresponds to the model shown in
    \figref{7025}, in which the minima and saddle-points have been
    incorrectly swapped.  The middle corresponds to \figref{7022},
    where the image parities were correct.  The lower panel
    corresponds to another model, where the image parities were
    correct but the time-ordering was incorrect.}
\label{fig:kapenc_compare_faulty}
\end{figure}

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\subsection{Test of mass-profile recovery} \label{sec:tests.t2}

The second test was to compare the mass distributions $\kappa(x,y)$ of
the sims and of the \spl models.  A visual comparison is presented for
the eight models in Figures~\ref{fig:6941}--\ref{fig:7022}, in the
lower-left versus lower-right panels.  We will summarize the mass
distributions drastically in a single number, the effective Einstein
radius.  Other measures for comparison of free-form lensing mass
distributions appear in \cite{2014arXiv1401.7990C}, but comparing
Einstein radii is already useful.

There is no standard way of defining the Einstein radius of a general
non-circular lens.  We adopt the simple definition
\begin{equation} \label{eq:effRE}
 \langle\kappa\rangle_{\Theta_{\rm E}} = 1,
\end{equation}
that is, the effective Einstein radius \ERf is such that the
mean $\kappa$ is unity inside a circle of radius \ERf centred at
the lens centre.

To illustrate, \figref{kapenc_compare_faulty} compares the circularly
averaged mass profiles three different models of one particular lens;
two of the models are shown in Figures~\ref{fig:7025} and
\ref{fig:7022}. Each panel in \figref{kapenc_compare_faulty} shows the
mean $\kappa$ within a circle of given radius.  The red curve is the
correct profile for the sim.  The two blue curves are the minimal and
maximal mean enclosed $\kappa$ from the internal ensemble in \spl.
Radial locations of the images are marked, along with the image
parities.  The region between the blue curves is shaded between the
radii of the innermost and the outermost images: this is the
confidence region from the modelling.  The definition \eqref{eq:effRE}
for \ERf corresponds to crossing the dashed horizontal line at $10^0$:
the red curve crosses the dashed line at the actual Einstein radius
\ERf[,act]; the recovered Einstein radius \ERf[,rec] and its
uncertainty are given by the blue curves crossing the dashed line.  We
see that in all three panels, the blue curves are shallower than the
red curve and \ERf[,rec] is more than \ERf[,act], by more than the
model uncertainties.  Now, steeper mass profiles tend to give wider
image separations --- recall that the image separation for a circular
isothermal lens is 2\ERf, whereas for a point mass it is more
\citep[see, e.g.,][]{2002LNP...608....1C} --- so \ERf[,rec] being too
high is really a consequence of the \spl models being too shallow for
the sims.

\begin{figure*}
  \centering
    \includegraphics[width=0.90\linewidth]{fig/eR_5}
  \caption{Model-recovered versus actual Einstein radii \ERf[,rec] and
    \ERf[,act].  Circles denote models where the image parities and
    time-orderings were correct, with filled circles for models by an
    expert.  Crosses mark models where the time-ordering was
    incorrect.  Plus signs indicate models flagged by the modeller as
    failures.}
  \label{fig:ER_per_sim}
\end{figure*}

\Figref{ER_per_sim} shows that \ERf[,rec] of the models tend
to be too high.  Discounting the models which were flagged by the
users as poor, the mean and standard deviation of the overestimate are
25\% and 20\%.  The models by an expert (PS) also show a similar bias,
having mean and standard deviation of the overestimate being 20\% and
8\%.  One of the reasons for this is that it is hard to get the center
of the lens on spot.  An offset leads to a flatter mass profile for
the model compared to the simulation.

% ====================================================================

\section{Outlook} \label{sec:todo}

This work has developed the concept of saddle-point contours in the
travel time of virtual photons, originally introduced by
\cite{1986ApJ...310..568B} for understanding image structure in strong
gravitational lenses, into a technique for mass-mapping lenses.
Despite being highly abstract, saddle-point contours look like
schematic arcs, and hence lend themselves to an intuitive markup tool
for lenses or lens candidates, which we call a spaghetti diagram.  At
the same time, saddle-point contours encode information about possible
mass distributions, which can be translated into input for an existing
lens-modelling engine \citep[GLASS, by][]{2014arXiv1401.7990C}.

\spl is an implementation of these ideas, enabling experienced but
non-professional lens enthusiasts to model newly-discovered lens
candidates from the \sw citizen-science platform.  The tests in this
paper indicate that such modelling would be both feasible and
scientifically interesting.

There is, however, plenty of room for improvement.

\begin{enumerate}
\item Currently, \spl does not attempt to model the source shape; the
  user identifies the brightest points on the image, and these are
  taken as images of a point-like source, whose positions must be
  reproduced exactly. For generating a synthetic image, a conical
  source profile is assumed. Fitting for the source profile to
  optimize resemblance to the observed lensed image after the lens
  model has been generated, is algorithmically straightforward
  \citep[cf.][]{2003ApJ...590..673W,2006MNRAS.371..983S} and planned
  to be implemented.  This would alleviate another problem with \spl,
  which is that there is no numerical figure of merit, and assessment
  of a model is a judgment call based on the synthetic image, and on
  whether the mass distribution and the arrival-time surface show
  suspect features.
\item \spl tends to overestimate the Einstein radius (evident from
  Figure \ref{fig:ER_per_sim}), and it is also apparent that the
  models tend to be too shallow.  The likely explanation is that while
  the sims are steeply peaked at the centre, the pixellated mass model
  fixes a comparatively large area near the central at constant
  density.  One way to solve this problem would be to introduce
  smaller pixels in the central region, thus enabled a steeper centre.
\item Another limitation so far in \spl is that the lens is assumed to
  be dominated by one galaxy, which puts most galaxy-group lenses
  beyond the reach of the modeler. Since complicated group lenses are
  some of the most interesting candidates present, removing this
  limitation is most desirable.  From the users' point of view, it
  would mean that spaghetti contours with more than one maximum can be
  allowed.  For examples, see Figure 5c in \citep{2001ApJ...557..594R}
  and Figure 4b in \cite{2003ApJ...590...39K}.
\item At present, a single false-color composite is used as the data.
  An option could be added to use all available filters, individually
  or in combination, at the user wishes.
\item As mentioned above, the option of revising an already-archived
  model is already available.  Desired now are tools for comparing
  different models of a given system, both visually and through
  different statistical measures.  As evidenced by a current
  collaborative modeling effort, a particularly interesting candidate
  can lead to an extended discussion and dozens of models.
\item Better tutorial materials are also needed, and this would
  address some of the problem areas found in the modeling challenge.
  For example, we saw in \secref{tests.t1} that volunteers are most
  prone to making errors in two situations: when in identifying an arc-
  like structure while placing the points, and in identifying the
  correct ordering of the points in nearly-symmetric configurations.
\end{enumerate}

The \spl program was developed by Kueng, with design suggestions from
Coles, Cornen and Saha, and feedback from all co-authors.  The
simulations were created by A.~More, in consultation with Marshall,
S.~More and Verma.  Modelling was done by Baeten, Cornen, McMillan,
Odermatt, Saha and Wilcox, with post-modelling analysis by Kueng and
Saha.  All authors participated in writing and editing the manuscript.

% ====================================================================

\section*{Acknowledgments}

We thank the Swiss Society for Astrophysics and Astronomy and the
Swiss Academy of Sciences. AV is supported by a research fellowship
from the Leverhulme Trust.

% ====================================================================

\appendix

\section{Relation to standard lensing formalism}\label{more-theory}

The description of the arrival-time surface in
Sections~\ref{sec:Fermat} and \ref{sec:arriv} omitted some details for
the sake of a more intuitive explanation.  The convergence $\kappa$
and the geometric time delay $t_{\rm geom}$ were left as
proportionalities (equations \ref{eq:kappa} and \ref{eq:tgeom}), and
the gravitational time delay $t_{\rm grav}$ was given in an implicit
form (equation \ref{eq:tgrav}).  Here we fill in the details.

The original formulation of the arrival-time surface appears in
equations (2.1) to (2.6) of \cite{1986ApJ...310..568B}.  Their
equations can be rearranged as follows.
\begin{equation} \label{cfBN86}
\begin{aligned}
t_{\rm geom} &= \frac{(1+z_L)}{2c} \frac{d_S}{d_L d_{LS}}
\left[ (x-s_x)^2 + (y-s_y)^2 \right] \\
\nabla^2 t_{\rm grav} &= -(1+z_L)\frac{8\pi G}{c^3} \, \Sigma(x,y) \\
\kappa(x,y) &= \frac{4\pi G}{c^2} \frac{d_L d_{LS}}{d_S}
               \times \Sigma(x,y)
\end{aligned}
\end{equation}
The symbols $d_L,d_S$ and $d_{LS}$ are angular-diameter distances,
respectively from observer to lens, observer to source, and lens to
source. We have replaced angular positions on the sky with positions
on the lens plane as
\begin{equation}
(x,y) = d_L (\theta_x,\theta_y) \,.
\end{equation}
In the concordance cosmology
\begin{equation}
d_{LS} = \frac{c}{H_0}\,\frac1{1+z_S} \,
\int_{z_L}^{z_S} \!\! \frac{dz}{\sqrt{\Omega_m(1+z)^3 + \Omega_\Lambda}}
%\raise.5ex\hbox{.}
\end{equation}
and similarly $d_L$ and $d_S$.

The first line of equation \eqref{cfBN86} is $t_{\rm geom}$ from
equation \eqref{eq:tgeom} with the proportionality filled in, and with
the source offset at $(s_x,s_y)$ rather than at the origin.

The last line of equation \eqref{cfBN86} fills in the proportionality
factor in equation \eqref{eq:kappa} for the convergence (or
dimensionless surface density) $\kappa$.

The middle line of equation \eqref{cfBN86} is a Poisson equation for
the gravitational time delay, and is equivalent to the implicit
expression \eqref{eq:tgrav}.  One way to verify the equivalence is to
consider the small circle in equation \eqref{eq:tgrav} as a region
where $\Sigma$ is constant, and approximate $\tgrav$ by its Taylor
expansion to $O(x^2,y^2)$.  Substituting in equations \eqref{cfBN86}
gives the Taylor coefficients in terms of $\Sigma$, and result
satisfies the expression \eqref{eq:tgrav}.  Alternatively, we can
proceed with a discrete form of the Poission equation from
\eqref{cfBN86}.  Discretising on a grid with spacing $\Delta$, we have
\begin{equation} \label{eq:relax}
\begin{aligned}
\tgrav(x,y) &=
\frac14 \Big[ \tgrav(x+\Delta,y) + \tgrav(x-\Delta,y) + {}\\
&\kern24pt    \tgrav(x,y+\Delta) + \tgrav(x,y-\Delta) \Big] \\
&+ (1+z_L)\frac{2G}{c^3} \, \pi\Delta^2 \, \Sigma(x,y) \,.
\end{aligned}
\end{equation}
This is recognisable as a formula for solving the two-dimensional
Poisson equation from \eqref{cfBN86} by relaxation.  Let us now
replace the average over four neighbouring points by the circular
average $\left\langle \tgrav(x_\subcirc,y_\subcirc) \right\rangle$ and
replace $\pi\Delta^2\,\Sigma(x,y)$ by the enclosed mass
$M(x_\subbullet,y_\subbullet)$.  These replacements are valid in the
limit of a small grid.  The result is the implicit
equation \eqref{eq:tgrav}.

% ====================================================================

\bibliographystyle{mn2e}
\bibliography{bib/ms}

% ====================================================================

\end{document}

% ====================================================================